Data Preparation

### 2019-01-11 Data Preparation.Rmd
Copy Paste

```{r libraries, message = FALSE}
library(tidyverse)
library(tidytext)

library(gsheet) # import spreadsheets from google drive
library(tm)
library(hunspell) # Stemmer
#library(rlist)
```

@@

## Setup
Before importing the spreadsheets from the drive, let us create two vectors `list_assessment` and `list_AoD` which respectively list the `19` types of assessment and the `18` aims of the degree (AoD) of the degree which we take into consideration in the analysis. The information comes from a spreadsheet from our google drive *DARS* which we save as `lists`. In addition, we also create two additional vectors `list_assessment_plot` and `list_AoD_plot` which contain a selection of prominent assessment types and AoD that we will consider when ploting the data[^1].
```{r import list}
lists_brut <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1239912347')

lists <- list(
  Assessment      = lists_brut$Assessment,
  Assessment_plot = lists_brut$Assessment[lists_brut$`Assessment for Plot` == 1],
  AoD             = lists_brut$Aim[!is.na(lists_brut$Aim)],
  AoD_plot        = lists_brut$Aim[lists_brut$`Aim for Plot`[!is.na(lists_brut$Aim)] == 1]
  )

rm(lists_brut)
```

@@

```{r import d_course d_assessment and d_ILO, cache = TRUE}
d_course <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1655700848') %>%
  mutate(`Missing from ILO File` = as.logical(`Missing from ILO File`))
print(d_course)

d_assessment <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1102665750') %>%
  select(- `Comment on Assessment`) %>%
  gather(Assessment, assessment_covered, - Code, factor_key = T) %>%
  mutate(assessment_covered = as.logical(assessment_covered)) %>%
  filter(assessment_covered) %>%
  select(- assessment_covered) %>%
  arrange(Code)
print(d_assessment)

d_ILO <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1896457050') %>%
  select(- c(Comments, Comments_for_Jeroen)) %>%
  gather(AoD, AoD_covered, -c(Code, Objectives), factor_key = T) %>%
  mutate(AoD_covered = as.logical(AoD_covered)) %>%
  rename(ILO = Objectives) %>%
  filter(AoD_covered) %>%
  select(- AoD_covered) %>%
  arrange(Code)
print(d_ILO)
```

@@

In order to determine which AoD each course covers with its ILOs, we first eliminate the AoD that are not covered by the ILOs, and then we keep only one instance of each combination of course and AoD in case a course had several ILOs covering the same AoD.
```{r d_AoD_ILO}
d_AoD_ILO <- d_ILO %>%
  select(Code, AoD) %>%
  unique

print(d_AoD_ILO)
```

@@

Now that we have the matrix `map_assessment_AoD`, we want to find out which AoDs are covered by a course through its assessments. To do this we create an empty tibble `d_AoD_assessment` to store which assessment is covered by each course, and which AoD said assessment covers.Then, we fill in the information with a loop.
In the loop, we first extract a row of `d_assessment` and save it as `observation`. Then, we determine the corresponding assessment which we save as `assessment`. Then, we use the matrix `map_assessment_AoD` to determine which AoD `assessment` covers and use `cbind` and `rbind` to add the information to the tibble `d_AoD_assessment`.
```{r d_AoD_assessment}
d_AoD_assessment <- tibble(Code        = character(0),
                           Assessment  = character(0),
                           AoD         = character(0))

for(i in 1 : nrow(d_assessment)){

@@

  if(length(AoD) >= 1) d_AoD_assessment <- rbind(d_AoD_assessment,
                                                 as.tibble(cbind(observation,
                                                                 AoD)))

}

rm(map_assessment_AoD, observation, AoD, AoD_covered, assessments, i)
```

@@

#### Combining AoD from ILOs and from assessments
Finally, we can use a `rbind` to combine the two tibbles `d_AoD_ILO` and `d_AoD_assessment`. We also use `unique` in case a course covers an AoD with both its ILOS and its assessments.
```{r rbind d_AoD_ILO and d_AoD_assessment}
d_AoD <- rbind(d_AoD_ILO,
               d_AoD_assessment) %>%
  distinct %>%
  arrange(Code)

rm(d_AoD_ILO, d_AoD_assessment)
print(d_AoD)
```

@@

Now that we have a clear overview of the distribution of AoDs (`d_AoD`) and assessments (`d_assessment`) among the courses, let us add variables to the tibble `d_course` that contain the information at the course level. For this we create the following three tibbles:
1) `d_ILO_detail` contains two columns indicating the code of the course and the number of ILOs it contains.
2) `d_assessment_detail` contains three columns indicating the code of the course, the number of assessments it covers and a list of the assessments it covers. 
3) `d_AoD_detail` contains three columns indicating the code of the course, the number of AoD it covers and a list of the AoD it covers. 
Then we use a `full_join` to add these variables to the tibble `d_course`.
```{r d_course detail}
d_ILO_detail <- d_ILO %>% 
  select(Code, ILO) %>%
  distinct() %>%
  count(Code) %>%
  rename(n_ILO = n)

d_assessment_detail <- d_assessment %>%
  group_by(Code) %>%
  summarise(n_assessment = n(),
            `Assessments Covered` = paste(Assessment, collapse = ", "))

d_AoD_detail <- d_AoD %>%
  group_by(Code) %>%
  summarise(n_AoD = n(),
            `AoD Covered` = paste(AoD, collapse = ", "))

d_course_detail <- d_ILO_detail %>%
  full_join(d_assessment_detail, by = "Code") %>%
  full_join(d_AoD_detail       , by = "Code")

rm(d_ILO_detail, d_assessment_detail, d_AoD_detail)
print(d_course_detail)
```

```{r join d_course_detail}
d_course <- full_join(d_course, d_course_detail, by = "Code") %>%
  select(Code, `Course Title`, Cluster, n_ILO, n_assessment, `Assessments Covered`, n_AoD, `AoD Covered`, everything())

rm(d_course_detail)
print(d_course)
```

The undegraduate research courses (*UGR-*) are only present at the `3000` level (advanced level). Yet, these course are also offered at the `2000` level (intermediate level). We use an `rbind` to duplicate the rows of the course `UGR3000` and mutate their `Code` to `UGR2000`.
```{r create UGR2000}
d_course <- d_course %>%
  rbind(filter(., Code %in% c("UGR3001", "UGR3002", "UGR3003", "UGR3005")) %>%
          mutate(Code = case_when(Code == "UGR3001" ~ "UGR2001",
                                  Code == "UGR3002" ~ "UGR2002",
                                  Code == "UGR3003" ~ "UGR2003",
                                  Code == "UGR3005" ~ "UGR2005")))

print(filter(d_course, Code %in% c("UGR2001", "UGR2002", "UGR2003", "UGR2005",
                                   "UGR3001", "UGR3002", "UGR3003", "UGR3005")))
```

@@

```{r extracting overview, cache = TRUE}
# Setup
n_catalogue <- length(corpus_catalogues)
calendar_years <- c("2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019")
d_description <- tibble(Code            = character(0),
                        `Calendar Year` = character(0),
                        Overview        = character(0))

course_code <- c("COR", "HUM", "SCI", "SSC", "SKI", "PRO", "UGR", "CAP") %>%
  paste(collapse = "|")

content_to_exclude <- c("^Core Courses \\(COR\\)",
                        "^Humanities \\(HUM\\)",
                        "^Sciences \\(SCI\\)",
                        "^Social Sciences \\(SSC\\)",
                        "^Skills Trainings \\(SKI\\)",
                        "^Projects \\(PRO\\)",
                        "^Undergraduate Research \\(UGR\\)",
                        "Appendix") %>%
  paste(collapse = "|")

# Loop
for(n in 1 : n_catalogue){

@@

      if(following_page %in% c(first_pages_overview, pages_to_exclude)) course_overview <- current_page_content
      else course_overview <- paste(current_page_content, following_page_content)

      # Save Overview
      Code <- substring(course_overview, first = 1, last = 7)      
      year <- calendar_years[n]
      d_description <- add_row(d_description,
                               Code            = Code,
                               `Calendar Year` = year,
                               Overview        = course_overview)
    } # close if-statement
  } # close for-loop (page)
} # close for-loop (n)

rm(corpus_catalogues,
   n_catalogue, calendar_years, course_code, content_to_exclude,
   n, cat, pages_to_exclude, first_three_letters, first_pages_overview, page_start_loop, page_end_loop,
   page, current_page, following_page, current_page_content, following_page_content,
   course_overview, Code, year)
print(d_description)
```

@@

```{r extracting description, cache = TRUE}
# Setup
symbol_to_line <- c("\r\n", "\n") %>%
  paste(collapse = "|")

start_descrip <- c("^Description", "^Course Description") %>%
  paste(collapse = "|")

end_descrip   <- c("^Literature$", "^Recommended Literature", "^ Required Litera",
                   "^Required Litera", "^ Literature$", "Literature \\(all") %>%
  paste(collapse = "|")

d_description <- d_description %>%
  mutate(Description = NA)

# Loop
for(course in 1 : nrow(d_description)){

@@

  overview_by_line <- strsplit(x = toString(overview), split = symbol_to_line)[[1]]
  line_start    <- grep(pattern = start_descrip, x = overview_by_line) + 1
  line_end      <- grep(pattern = end_descrip  , x = overview_by_line) - 1

  # for overviews w/o Literature section, find section Instructional Format.
  if(is_empty(line_end)){
    line_end <- grep(pattern = "^Instructional format$", x = overview_by_line) - 1
    print(paste("line_end irregular for", d_description$Code[course],
                "in catalogue", d_description$`Calendar Year`[course]))
    } # close if-statement

@@

}

rm(symbol_to_line, start_descrip, end_descrip,
   course, overview, overview_by_line, line_start, line_end, description)
print(d_description)
```

@@

Since we want to use tools from dplyr and tidytext we must first transform our corpus into a dataframe.
```{r d_manual}
d_manual <- tidy(corpus_manuals) %>%
  select(id, text) %>%
  mutate(Code = substring(id, 1, 7)) %>%
  select(-id) %>%
  rename(Manual = text)

rm(corpus_manuals)
print(d_manual)
```

@@

To put everything into tidy text format, we first create three tibbles `d_overview_tidy`, `d_description_tidy`, `d_manual` that respectively store the overviews, descriptions and text from manuals in the tidy text format, with one row per course-year-word (and course-word for `d_manual`).
```{r tidy d_description}
d_overview <- d_description %>%
  select(Code, `Calendar Year`, Overview) %>%
  unnest_tokens(output = word, input = Overview)

d_description <- d_description %>%
  select(Code, `Calendar Year`, Description) %>%
  unnest_tokens(output = word, input = Description)

d_manual <- unnest_tokens(d_manual, output = word, input = Manual)
```

### Stemming
We then use the function `hunspell__stem`, which returns valid stems for a given word, to stem the words in the tibbles `d_overview_tidy` and `d_description_tidy`. We first create a function `stem_hunspell` which, given a term, returns its most basic stem (the last one from the list of stem returned by `hunspell__stem`). We then apply `stem_hunspell` on the words of `d_overview_tidy`. Since, `stem_hunspell` is not a vectorized function and the number of words in `d_overview_tidy` is large, we first use distinct on the `word` variable to find a list containing *once* each term present in the course overviews. Then, we apply the function `stem_hunspell` on this list. We save the terms and their stems in the tibble `stems`. Since the course descriptions can be conceptualized as a subset of the course overviews, the tibble `stems` contains all the terms present in the descriptions and can therefore safely used to stem the words in `d_description_tidy`.
```{r stemming dictionary, cache = TRUE}
stem_hunspell <- function(term) {
  stems   <- hunspell_stem(term)[[1]] # look up the term in the dictionary
  n_stems <- length(stems)            # number of stems
  if (n_stems == 0) term              # if no stems, return original term
  else              stems[[n_stems]]  # if multiple stems, return last (most basic) stem
}

dictionary <- select(d_overview, Code, word) %>%
  rbind(d_manual) %>% 
  select(word) %>%
  distinct %>%
  mutate(word_stem = purrr::map_chr(.x = word,
                                    .f = stem_hunspell))

# terms for which the stem differs from the original word
filter(dictionary, word != word_stem)
```

@@

d_description <- stem_with_dictionary(d_description)
d_overview <- stem_with_dictionary(d_overview)
d_manual <- stem_with_dictionary(d_manual)

rm(stems, stem_hunspell)
print(d_overview_stem) # See humanities - humanity
```
### Removing Stopwords
```{r}

# Own stop words
sw_own <- c("______________________________________________________________________________")

remove_sw <- function(data) data %>%
  filter(! word %in% c(stop_words$word, sw_own))

d_description <- remove_sw(d_description)
d_overview <- remove_sw(d_overview)
d_manual <- remove_sw(d_manual)
```

@@

Each faculty had a different code for their calendars, we want to standardise them so we can compare. We convert everything into UCM calendar codes (the correspondance between calendars was given to us by R. Vos) and save it under a variable `Period`:
```{r}
d_transcript <- d_transcript %>%
  mutate(Period = case_when(Period == 1 ~ "1 to 6"  ,
                            Period == 2 ~ "1 to 3"  ,
                            Period == 3 ~ "4 to 6"  ,
                            Period == 680 ~ "5"     , # exception
                            Period == 350 ~ "3 to 5", # exception
                            Period == 403 ~ "4 to 5", # exception
                            Period == 900 ~ "1 to 6", # exception
                            between(Period, 100, 199) ~ "1",
                            between(Period, 200, 299) ~ "2",
                            between(Period, 300, 399) ~ "3",
                            between(Period, 400, 499) ~ "4",
                            between(Period, 500, 599) ~ "5",
                            between(Period, 600, 660) ~ "6",
                            between(Period, 901, 999) ~ "1"))

```

@@

First, lets find out, which year was a student's first year and which year was a student's last year (we will also create a variable `Subs_year` that we will substract later to transform from Calendar Year to UCM Year, and a second variable `LengthUCM` that shows us how many years each student stayed at the college):
```{r, eval = FALSE}

d_year <- d_grade %>%
  select(ID, Code, Year, Period, Grade,`Booking Status Description`, `Course concentration`) %>%
  group_by(ID) %>%
  summarise(FirstYear=min(Year), LastYear=max(Year)) %>%
  mutate(LengthUCM = LastYear-FirstYear, Subs_year=FirstYear-1)
```

@@

```{r, eval = FALSE}

d_year1 <- d_grade %>%
  select(ID, Code, Year, Period, Grade, `Booking Status Description`, `Course concentration`) %>%  
  group_by( ID, Year) %>%
  summarise(Count = n())%>%
  mutate(Present = case_when(Count > 0 ~ Year))%>%
  select(-Count) %>%
  spread(Year,Present, fill = 0)

```

@@

```{r}

d_time <- left_join(d_year1, d_year)

```

@@

```{r, eval = FALSE}
var_years <- as.character(2007:2017)

d_time <- d_time %>% 
  mutate_at(
    vars(var_years),
    funs(. - Subs_year)) %>%
  mutate_at(
    vars(var_years),
    funs(case_when(. < 0 ~ 9999999999, #HORRIBLE FIX
                   T     ~ .)))

d_time[d_time == 9999999999] <- NA

```

@@

```{r, eval = FALSE}
d_students <- d_grade %>%
  select(ID, Code, Year, Period, Grade, Attempt,`Booking Status Description`, `Course concentration`) %>%
  arrange(ID, Year, Period)%>%
  left_join(d_student_concentration, by="ID") %>%
  mutate(Grade = as.numeric( sub(",",".", Grade))) %>%
  replace_na(list(Grade=-1))

# Years dictionary:
year_dic <- d_time %>%
  select(-FirstYear,-LastYear,-LengthUCM,-Subs_year)%>%
  gather(Cal_Year,UCM_Year,var_years)%>%
  drop_na(.) %>%
  mutate(Year=as.integer(Cal_Year))%>%
  select(-Cal_Year)

#adding the ucm year to transcripts:
d_students <- d_students %>%
  left_join(year_dic, by=c("ID","Year"))

#removing repeated values:
d_students <- d_students[!duplicated(d_students),]

```

@@

For this we keep only the capstones. We expect some people to have repeated capstones:
```{r, eval = FALSE}
capstone_count <- d_students %>%
  filter(Code=="CAP3000") %>%
  group_by(ID)%>%
  summarise(Count=n())

capstone_count
```
However, what is surptising is that some people have two different passed capstones:

```{r, eval = FALSE}
passed_capstone_count <- d_students %>%
  filter(Code=="CAP3000", `Booking Status Description`=="Completed with Success") %>%
  group_by(ID) %>%
  summarise(Count=n(), Max = max(Grade), Min = min(Grade), Attempts= max(Attempt))

```

@@

First we get the years for which there is a corresponding passing capstone grade, and check for repetition:
```{r, eval = FALSE}

capstone_finish_year <- inner_join(passed_capstone_grades, d_students)

head(capstone_finish_year)

count_before <- passed_capstone_grades %>%
  group_by(ID) %>%
  summarise(counts_of_id= n())

count_after <- capstone_finish_year %>%
  group_by(ID) %>%
  summarise(counts_of_id= n())%>%
  filter(counts_of_id > 1)  #these students have repeated capstones.

count_after

```
From count_after we see that student 544450 has two passed capstones at different moments in time, we already knew this and we must ask Richard why this is the case. However, since both capstones where passed in 2010, this creates no problem for our current purposes. However, let's remove the later version.
We will also rename `Year` to `Capstone Year` and keep only the columns `Capstone Year` and `ID`
```{r}
to_remove <- capstone_finish_year%>%
  filter(ID==544450)
to_remove <- to_remove[duplicated(to_remove$ID),]

>>>>

```{r libraries, message = FALSE}
library(tidyverse)
library(tidytext) 
library(gsheet) # import spreadsheets from google drive
library(tm)
library(hunspell) # Stemmer
#library(rlist)
```

@@

## Setup
Before importing the spreadsheets from the drive, let us create two vectors `list_assessment` and `list_AoD` which respectively list the `19` types of assessment and the `18` aims of the degree (AoD) of the degree which we take into consideration in the analysis. The information comes from a spreadsheet from our google drive *DARS* which we save as `lists`. In addition, we also create two additional vectors `list_assessment_plot` and `list_AoD_plot` which contain a selection of prominent assessment types and AoD that we will consider when ploting the data[^1].
```{r import list}
lists_brut <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1239912347')
lists <- list(
  Assessment      = lists_brut$Assessment,
  Assessment_plot = lists_brut$Assessment[lists_brut$`Assessment for Plot` == 1],
  AoD             = lists_brut$Aim[!is.na(lists_brut$Aim)],
  AoD_plot        = lists_brut$Aim[lists_brut$`Aim for Plot`[!is.na(lists_brut$Aim)] == 1]
  )
rm(lists_brut)
```

@@

```{r import d_course d_assessment and d_ILO, cache = TRUE}
d_course <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1655700848') %>%
  mutate(`Missing from ILO File` = as.logical(`Missing from ILO File`))
print(d_course)
d_assessment <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1102665750') %>%
  select(- `Comment on Assessment`) %>%
  gather(Assessment, assessment_covered, - Code, factor_key = T) %>%
  mutate(assessment_covered = as.logical(assessment_covered)) %>%
  filter(assessment_covered) %>%
  select(- assessment_covered) %>%
  arrange(Code)
print(d_assessment)
d_ILO <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1896457050') %>%
  select(- c(Comments, Comments_for_Jeroen)) %>%
  gather(AoD, AoD_covered, -c(Code, Objectives), factor_key = T) %>%
  mutate(AoD_covered = as.logical(AoD_covered)) %>%
  rename(ILO = Objectives) %>%
  filter(AoD_covered) %>%
  select(- AoD_covered) %>%
  arrange(Code)
print(d_ILO)
```

@@

In order to determine which AoD each course covers with its ILOs, we first eliminate the AoD that are not covered by the ILOs, and then we keep only one instance of each combination of course and AoD in case a course had several ILOs covering the same AoD.
```{r d_AoD_ILO}
d_AoD_ILO <- d_ILO %>%
  select(Code, AoD) %>%
  unique
print(d_AoD_ILO)
```

@@

Now that we have the matrix `map_assessment_AoD`, we want to find out which AoDs are covered by a course through its assessments. To do this we create an empty tibble `d_AoD_assessment` to store which assessment is covered by each course, and which AoD said assessment covers.Then, we fill in the information with a loop.
In the loop, we first extract a row of `d_assessment` and save it as `observation`. Then, we determine the corresponding assessment which we save as `assessment`. Then, we use the matrix `map_assessment_AoD` to determine which AoD `assessment` covers and use `cbind` and `rbind` to add the information to the tibble `d_AoD_assessment`.
```{r d_AoD_assessment}
d_AoD_assessment <- tibble(Code        = character(0),
                           Assessment  = character(0),
                           AoD         = character(0))
for(i in 1 : nrow(d_assessment)){

@@

  if(length(AoD) >= 1) d_AoD_assessment <- rbind(d_AoD_assessment,
                                                 as.tibble(cbind(observation,
                                                                 AoD)))
}
rm(map_assessment_AoD, observation, AoD, AoD_covered, assessments, i)
```

@@

#### Combining AoD from ILOs and from assessments
Finally, we can use a `rbind` to combine the two tibbles `d_AoD_ILO` and `d_AoD_assessment`. We also use `unique` in case a course covers an AoD with both its ILOS and its assessments.
```{r rbind d_AoD_ILO and d_AoD_assessment}
d_AoD <- rbind(d_AoD_ILO,
               d_AoD_assessment) %>%
  distinct %>%
  arrange(Code)
rm(d_AoD_ILO, d_AoD_assessment)
print(d_AoD)
```

@@

Now that we have a clear overview of the distribution of AoDs (`d_AoD`) and assessments (`d_assessment`) among the courses, let us add variables to the tibble `d_course` that contain the information at the course level. For this we create the following three tibbles:
1) `d_ILO_detail` contains two columns indicating the code of the course and the number of ILOs it contains.
2) `d_assessment_detail` contains three columns indicating the code of the course, the number of assessments it covers and a list of the assessments it covers. 
3) `d_AoD_detail` contains three columns indicating the code of the course, the number of AoD it covers and a list of the AoD it covers. 
Then we use a `full_join` to add these variables to the tibble `d_course`.
```{r d_course detail}
d_ILO_detail <- d_ILO %>% 
  select(Code, ILO) %>%
  distinct() %>%
  count(Code) %>%
  rename(n_ILO = n)
d_assessment_detail <- d_assessment %>%
  group_by(Code) %>%
  summarise(n_assessment = n(),
            `Assessments Covered` = paste(Assessment, collapse = ", "))
d_AoD_detail <- d_AoD %>%
  group_by(Code) %>%
  summarise(n_AoD = n(),
            `AoD Covered` = paste(AoD, collapse = ", "))
d_course_detail <- d_ILO_detail %>%
  full_join(d_assessment_detail, by = "Code") %>%
  full_join(d_AoD_detail       , by = "Code")
rm(d_ILO_detail, d_assessment_detail, d_AoD_detail)
print(d_course_detail)
```

```{r join d_course_detail}
d_course <- full_join(d_course, d_course_detail, by = "Code") %>%
  select(Code, `Course Title`, Cluster, n_ILO, n_assessment, `Assessments Covered`, n_AoD, `AoD Covered`, everything())
rm(d_course_detail)
print(d_course)
```

The undegraduate research courses (*UGR-*) are only present at the `3000` level (advanced level). Yet, these course are also offered at the `2000` level (intermediate level). We use an `rbind` to duplicate the rows of the course `UGR3000` and mutate their `Code` to `UGR2000`.
```{r create UGR2000}
d_course <- d_course %>%
  rbind(filter(., Code %in% c("UGR3001", "UGR3002", "UGR3003", "UGR3005")) %>%
          mutate(Code = case_when(Code == "UGR3001" ~ "UGR2001",
                                  Code == "UGR3002" ~ "UGR2002",
                                  Code == "UGR3003" ~ "UGR2003",
                                  Code == "UGR3005" ~ "UGR2005")))
print(filter(d_course, Code %in% c("UGR2001", "UGR2002", "UGR2003", "UGR2005",
                                   "UGR3001", "UGR3002", "UGR3003", "UGR3005")))
```

@@

```{r extracting overview, cache = TRUE}
# Setup
n_catalogue <- length(corpus_catalogues)
calendar_years <- c("2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019")
d_description <- tibble(Code            = character(0),
                        `Calendar Year` = character(0),
                        Overview        = character(0))
course_code <- c("COR", "HUM", "SCI", "SSC", "SKI", "PRO", "UGR", "CAP") %>%
  paste(collapse = "|")
content_to_exclude <- c("^Core Courses \\(COR\\)",
                        "^Humanities \\(HUM\\)",
                        "^Sciences \\(SCI\\)",
                        "^Social Sciences \\(SSC\\)",
                        "^Skills Trainings \\(SKI\\)",
                        "^Projects \\(PRO\\)",
                        "^Undergraduate Research \\(UGR\\)",
                        "Appendix") %>%
  paste(collapse = "|")
# Loop
for(n in 1 : n_catalogue){

@@

      if(following_page %in% c(first_pages_overview, pages_to_exclude)) course_overview <- current_page_content
      else course_overview <- paste(current_page_content, following_page_content)
      # Save Overview
      Code <- substring(course_overview, first = 1, last = 7)      
      year <- calendar_years[n]
      d_description <- add_row(d_description,
                               Code            = Code,
                               `Calendar Year` = year,
                               Overview        = course_overview)
    } # close if-statement
  } # close for-loop (page)
} # close for-loop (n)
rm(corpus_catalogues,
   n_catalogue, calendar_years, course_code, content_to_exclude,
   n, cat, pages_to_exclude, first_three_letters, first_pages_overview, page_start_loop, page_end_loop,
   page, current_page, following_page, current_page_content, following_page_content,
   course_overview, Code, year)
print(d_description)
```

@@

```{r extracting description, cache = TRUE}
# Setup
symbol_to_line <- c("\r\n", "\n") %>%
  paste(collapse = "|")
start_descrip <- c("^Description", "^Course Description") %>%
  paste(collapse = "|")
end_descrip   <- c("^Literature$", "^Recommended Literature", "^ Required Litera",
                   "^Required Litera", "^ Literature$", "Literature \\(all") %>%
  paste(collapse = "|")
d_description <- d_description %>%
  mutate(Description = NA)
# Loop
for(course in 1 : nrow(d_description)){

@@

  overview_by_line <- strsplit(x = toString(overview), split = symbol_to_line)[[1]]
  line_start    <- grep(pattern = start_descrip, x = overview_by_line) + 1
  line_end      <- grep(pattern = end_descrip  , x = overview_by_line) - 1
  # for overviews w/o Literature section, find section Instructional Format.
  if(is_empty(line_end)){
    line_end <- grep(pattern = "^Instructional format$", x = overview_by_line) - 1
    print(paste("line_end irregular for", d_description$Code[course],
                "in catalogue", d_description$`Calendar Year`[course]))
    } # close if-statement

@@

}
rm(symbol_to_line, start_descrip, end_descrip,
   course, overview, overview_by_line, line_start, line_end, description)
print(d_description)
```

@@

Since we want to use tools from dplyr and tidytext we must first transform our corpus into a dataframe.
```{r d_manual}
d_manual <- tidy(corpus_manuals) %>%
  select(id, text) %>%
  mutate(Code = substring(id, 1, 7)) %>%
  select(-id) %>%
  rename(Manual = text)
rm(corpus_manuals)
print(d_manual)
```

@@

To put everything into tidy text format, we first create three tibbles `d_overview_tidy`, `d_description_tidy`, `d_manual` that respectively store the overviews, descriptions and text from manuals in the tidy text format, with one row per course-year-word (and course-word for `d_manual`).
```{r tidy d_description}
d_overview <- d_description %>%
  select(Code, `Calendar Year`, Overview) %>%
  unnest_tokens(output = word, input = Overview)
d_description <- d_description %>%
  select(Code, `Calendar Year`, Description) %>%
  unnest_tokens(output = word, input = Description)
d_manual <- unnest_tokens(d_manual, output = word, input = Manual)
```

### Stemming
We then use the function `hunspell__stem`, which returns valid stems for a given word, to stem the words in the tibbles `d_overview_tidy` and `d_description_tidy`. We first create a function `stem_hunspell` which, given a term, returns its most basic stem (the last one from the list of stem returned by `hunspell__stem`). We then apply `stem_hunspell` on the words of `d_overview_tidy`. Since, `stem_hunspell` is not a vectorized function and the number of words in `d_overview_tidy` is large, we first use distinct on the `word` variable to find a list containing *once* each term present in the course overviews. Then, we apply the function `stem_hunspell` on this list. We save the terms and their stems in the tibble `stems`. Since the course descriptions can be conceptualized as a subset of the course overviews, the tibble `stems` contains all the terms present in the descriptions and can therefore safely used to stem the words in `d_description_tidy`.
```{r stemming dictionary, cache = TRUE}
stem_hunspell <- function(term) {
  stems   <- hunspell_stem(term)[[1]] # look up the term in the dictionary
  n_stems <- length(stems)            # number of stems
  if (n_stems == 0) term              # if no stems, return original term
  else              stems[[n_stems]]  # if multiple stems, return last (most basic) stem
}
dictionary <- select(d_overview, Code, word) %>%
  rbind(d_manual) %>% 
  select(word) %>%
  distinct %>%
  mutate(word_stem = purrr::map_chr(.x = word,
                                    .f = stem_hunspell))
# terms for which the stem differs from the original word
filter(dictionary, word != word_stem)
```

@@

d_description <- stem_with_dictionary(d_description)
d_overview <- stem_with_dictionary(d_overview)
d_manual <- stem_with_dictionary(d_manual)
rm(stems, stem_hunspell, stem_with_dictionary)
print(d_overview) # See humanities - humanity
```
### Removing Stopwords
```{r}
# Own stop words
sw_own <- c("______________________________________________________________________________")
remove_sw <- function(data) data %>%
  filter(! word %in% c(stop_words$word, sw_own))
d_description <- remove_sw(d_description)
d_overview <- remove_sw(d_overview)
d_manual <- remove_sw(d_manual)
rm(sw_own, remove_sw)
```

@@

Each faculty had a different code for their calendars, we want to standardise them so we can compare. We convert everything into UCM calendar codes (the correspondance between calendars was given to us by R. Vos) and save it under a variable `Period`:
```{r}
d_transcript <- d_transcript %>%
  mutate(Period = case_when(Period == 1 ~ "1 to 6"  ,
                            Period == 2 ~ "1 to 3"  ,
                            Period == 3 ~ "4 to 6"  ,
                            Period == 680 ~ "5"     , # exception
                            Period == 350 ~ "3 to 5", # exception
                            Period == 403 ~ "4 to 5", # exception
                            Period == 900 ~ "1 to 6", # exception
                            between(Period, 100, 199) ~ "1",
                            between(Period, 200, 299) ~ "2",
                            between(Period, 300, 399) ~ "3",
                            between(Period, 400, 499) ~ "4",
                            between(Period, 500, 599) ~ "5",
                            between(Period, 600, 660) ~ "6",
                            between(Period, 901, 999) ~ "1"))
```

@@

First, lets find out, which year was a student's first year and which year was a student's last year (we will also create a variable `Subs_year` that we will substract later to transform from Calendar Year to UCM Year, and a second variable `LengthUCM` that shows us how many years each student stayed at the college):
```{r, eval = FALSE}
d_year <- d_grade %>%
  select(ID, Code, Year, Period, Grade,`Booking Status Description`, `Course concentration`) %>%
  group_by(ID) %>%
  summarise(FirstYear=min(Year), LastYear=max(Year)) %>%
  mutate(LengthUCM = LastYear-FirstYear, Subs_year=FirstYear-1)
```

@@

```{r, eval = FALSE}
d_year1 <- d_grade %>%
  select(ID, Code, Year, Period, Grade, `Booking Status Description`, `Course concentration`) %>%  
  group_by( ID, Year) %>%
  summarise(Count = n())%>%
  mutate(Present = case_when(Count > 0 ~ Year))%>%
  select(-Count) %>%
  spread(Year,Present, fill = 0)
```

@@

```{r, eval = FALSE}
d_time <- left_join(d_year1, d_year)
```

@@

```{r, eval = FALSE}
var_years <- as.character(2007:2017)
d_time <- d_time %>% 
  mutate_at(
    vars(var_years),
    funs(. - Subs_year)) %>%
  mutate_at(
    vars(var_years),
    funs(case_when(. < 0 ~ 9999999999, #HORRIBLE FIX
                   T     ~ .)))
d_time[d_time == 9999999999] <- NA
```

@@

```{r, eval = FALSE}
d_students <- d_grade %>%
  select(ID, Code, Year, Period, Grade, Attempt,`Booking Status Description`, `Course concentration`) %>%
  arrange(ID, Year, Period)%>%
  left_join(d_student_concentration, by="ID") %>%
  mutate(Grade = as.numeric( sub(",",".", Grade))) %>%
  replace_na(list(Grade=-1))
# Years dictionary:
year_dic <- d_time %>%
  select(-FirstYear,-LastYear,-LengthUCM,-Subs_year)%>%
  gather(Cal_Year,UCM_Year,var_years)%>%
  drop_na(.) %>%
  mutate(Year=as.integer(Cal_Year))%>%
  select(-Cal_Year)
#adding the ucm year to transcripts:
d_students <- d_students %>%
  left_join(year_dic, by=c("ID","Year"))
#removing repeated values:
d_students <- d_students[!duplicated(d_students),]
```

@@

For this we keep only the capstones. We expect some people to have repeated capstones:
```{r, eval = FALSE}
capstone_count <- d_students %>%
  filter(Code=="CAP3000") %>%
  group_by(ID)%>%
  summarise(Count=n())
capstone_count
```
However, what is surptising is that some people have two different passed capstones:

```{r, eval = FALSE}
passed_capstone_count <- d_students %>%
  filter(Code=="CAP3000", `Booking Status Description`=="Completed with Success") %>%
  group_by(ID) %>%
  summarise(Count=n(), Max = max(Grade), Min = min(Grade), Attempts= max(Attempt))
```

@@

First we get the years for which there is a corresponding passing capstone grade, and check for repetition:
```{r, eval = FALSE}
capstone_finish_year <- inner_join(passed_capstone_grades, d_students) 
head(capstone_finish_year)
count_before <- passed_capstone_grades %>%
  group_by(ID) %>%
  summarise(counts_of_id= n())
count_after <- capstone_finish_year %>%
  group_by(ID) %>%
  summarise(counts_of_id= n())%>%
  filter(counts_of_id > 1)  #these students have repeated capstones.
count_after
